Simulation Tools for Small Area Estimation: Introducing the \proglang{R}-package \proglang{saeSim}

### 2014-10-10 guidelinesAJS.Rnw
saeSim paper template

%% almost as usual
\author{Matthias Templ\\ Vienna University of Technology \And 
        Second Author\\Plus Affiliation}
\title{A Capitalized Title: Something about a Topic}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Matthias Templ, Second Author} %% comma-separated
\Plaintitle{A Capitalized Title: Something about a Topic} %% without formatting
\Shorttitle{A Capitalized Title} %% a short title (if necessary)

@@

}
\Keywords{keywords, comma-separated, not capitalized, \proglang{R}}
\Plainkeywords{keywords, comma-separated, not capitalized, R} %% without formatting
%% at least one keyword must be supplied

@@

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Matthias Templ\\
  Department of Statistics and Probability Theory\\
  Vienna University of Technology\\
  A-1040 Vienna, Austria\\
  E-mail: \email{templ@tuwien.ac.at}\\
  URL: \url{http://www.statistik.tuwien.ac.at/public/templ}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

@@

<<echo=FALSE>>=
suppressPackageStartupMessages(library(saeSim))
@

<<>>=
sim_base_lm()
@

>>>>

%% almost as usual
\author{Sebastian Warnholz\\ Freie Universit\"at Berlin \And 
        Timo Schmid \\ Freie Universit\"at Berlin}
\title{Simulation Tools for Small Area Estimation: Introducing the \proglang{R}-package \proglang{saeSim}}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Sebastian Warnholz, Timo Schmid} %% comma-separated
\Plaintitle{Simulation Tools for Small Area Estimation: Introducing the R-package saeSim} %% without formatting
\Shorttitle{Simulation Tools for Small Area Estimation} %% a short title (if necessary)

@@

}
\Keywords{package, small area estimation, reproducible research, simulation, \proglang{R}}
\Plainkeywords{package, small area estimation, reproducible research, simulation, R} %% without formatting
%% at least one keyword must be supplied

@@

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Sebastian Warnholz\\
  Department of Economics\\
  Freie Universit\"at Berlin\\
  D-14195 Berlin, Germany\\
  E-mail: \email{Sebastian.Warnholz@fu-berlin.de}\\
  URL: \url{http://www.wiwiss.fu-berlin.de/fachbereich/vwl/Schmid/Team/Warnholz.html}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

@@

% Loading (and installing) necessary packages.
<<echo=FALSE>>=
if(!suppressPackageStartupMessages(require(saeSim))) {
  if(!require(devtools)) {
    install.packages("devtools", repos = "http://cran.at.r-project.org/", dependencies = TRUE)
    library(devtools)
  }

  install.packages(c("dplyr", "ggplot2", "MASS", "spdep"), 
                   repos = "http://cran.at.r-project.org/", dependencies = TRUE)
  install_github("wahani/saeSim")
}
@

### 2014-10-10 saeSim.Rnw
saeSim paper template

\section{Introduction}

Besides the general objectives of the Austrian Journal of
Statistics, this should contain general instructions for
preparation of manuscripts to be submitted. This \LaTeXe\ file can
also be used as a template for the setting up of a text to be
submitted in computer readable form. In any case, use \texttt{pdflatex} to compile your \LaTeX \ file.

The style file of the Austrian Journal of Statsitics uses mainly the style file from the Journal of Statistical Software (JSS) plus minor modifications. We thank Achim Zeileis for his great work on the JSS style file.

The Austrian Journal of Statistics is freely available over the net. Free and open-source tools used for production of your manuscript are therefore also highly supported. Manuscripts can also be sent as \texttt{Sweave} \citep[see, e.g.,][]{leisch02} or \texttt{knitr} \citep{yihui13} files.

@@

\section{Objectives of the journal}

The journal is published approximately quarterly by the Austrian
Statistical Society. Its general objective is to promote and
extend the use of statistical methods in all kind of theoretical
and applied disciplines. Special emphasis is on methods and
results in official or governmental statistics.

Original papers and review articles will be published in the
Austrian Journal of Statistics if judged consistently with these
general aims. All papers will be refereed.

Special issues will appear from time to time. Each
section will have as a theme a specialized area of statistical
application, theory, or methodology.

Technical notes or problems for considerations under Shorter
Communications are also invited.

One author of each article obtains a free copy of the entire issue
in which the article is contained. Members of the Austrian
Statistical Society receive a copy of the Journal free of charge.
Specific volumes may be purchased from the Society.

Articles will also be made available through the web at 
\begin{center}
\href{http://www.ajs.or.at}{http://www.ajs.or.at}
\end{center}

\subsection{Statistics: theory and applications}

The section for {\bf theory} and methods publishes articles that
make original contributions to the foundations, theoretical
development, and methodology of statistics and probability. {\em
Theory and Methods\/} should be interpreted broadly to include all
techniques relevant to statistics and probability. This may
include computational and graphical methods as well as more
traditional mathematical methods. The research reported should be
motivated by a scientific or practical problem and illustrated by
application of the proposed methodology to that problem.
Illustration of techniques with real data is especially welcomed
and strongly encouraged.
\medskip

{\bf Applications} and case studies are published if they present
original articles which
\begin{list}{$\bullet$}{\setlength{\topsep}{0mm}\setlength{\itemsep}{-1mm}}
 \item show analysis with real data that are statistically
       innovative as well as scientifically and practically relevant, or
 \item contribute substantially to a scientific field through the
       use of sound statistical methods, or
 \item present new and useful data, such as a new life table for
       a segment of the population or a new social or economic
       indicator, or
 \item using empirical tests, examine or illustrate for an important
       application the utility of a valuable statistical technique, or
 \item evaluate the quality of important data sources.
\end{list}
Note that careful analysis of data of substantive importance may be
published even there are no methodological innovations.

\subsection{Official statistics}

The section for {\bf official statistics} publishes articles which
deal with conceptual methods and computational issues. The
different fields covered include
\begin{list}{$\bullet$}{\setlength{\topsep}{0mm}\setlength{\itemsep}{-1mm}}
 \item population and social statistics
 \item economic statistics
 \item price statistics
 \item regional and urban statistics
 \item economic and social statistics
 \item tourism statistics
 \item research, development and innovation statistics
 \item environmental statistics including environmental accounting and
       environmental indicators
 \item national accounts including input-output tables and analysis.
\end{list}
Especially welcomed are articles which are concerned in
\begin{list}{$\bullet$}{\setlength{\topsep}{0mm}\setlength{\itemsep}{-1mm}}
 \item statistical disclosure control methods
 \item model-based methods
 \item publication policies (presentation of statistics)
 \item data pre-processing
 \item data visualisation
 \item survey methodology
\end{list}
Of course, papers treating national as well as international
concepts (e.g.~EU-statistics) are of interest.

\section{Style files for the manuscript}

The Journal is produced on the basis of computer readable
manuscripts provided by the authors.

In order to obtain a uniform layout and to ease the production as
much as possible it is highly recommended to use \LaTeXe\ for the
text processing. Graphics should be added in pdf or png format
 at suitable places of the manuscript.

We highly recommend
to follow exactly the provided instructions already at the very
beginning when writing the paper. The present document can be used
directly as template. The file named {\tt guidelinesAJS.tex} may be
obtained from the site

\href{http://www.ajs.or.at/index.php/ajs/about/submissions\#authorGuidelines}{http://www.ajs.or.at/index.php/ajs/about/submissions\#authorGuidelines}

Please consider title case for titles and for references. Sentence case for  sections and subsections.

\subsection{Submission}

A manuscript should be submitted to
\begin{center}
\href{http://www.ajs.or.at/index.php/ajs/about/submissions\#onlineSubmissions}{http://www.ajs.or.at/index.php/ajs/about/submissions\#onlineSubmissions}
\end{center}

in
electronic form following the editorial system. For further questions for submission and on the editorial system, send an e-mail to \href{templ@statistik.tuwien.ac.at}{templ@statistik.tuwien.ac.at}.

\subsection{Formal configuration}

Each article must have a summary in English and German which
inform the reader quickly on the contents and the substantial
results. The length surely depends on the contents, however, it
should not exceed 150 words. Mathematical formulas and references should be
avoided in the summary.

Headings and footings as well as pagination will be added during
the production. Footnotes should be avoided as much as possible.

The article should be structured by enumerated sections and
subsections. The title should be in title case style, sections and subsections in section case style. Graphics and tables must contain captions, above
tables and below graphics (figures). Examples are (see Figure~\ref{Fig3.4} and Table~\ref{Tab3.4})

\begin{figure}[hbt]
\centerline{\sl Here should be a graphic using the includegraphics command (pdf or png format).}
\caption{\label{Fig3.4}The caption of the figure to be filled in.} 
\end{figure}

\begin{table}[hbt]
\caption{\label{Tab3.4}80 measured values of access time.}
\vspace*{-5mm}
\small
\hspace{1.5cm}
\begin{center}
\begin{tabular}{llllllll}
70.0 & 68.6 & 67.9 & 66.3 & 71.0 & 64.2 & 69.6 & 71.0\\
... & ... & ... & ... & ... & ... & ... & ...\\
%69.7 & 69.0 & 73.4 & 69.0 & 70.1 & 69.8 & 69.0 & 73.0\\
%70.0 & 69.1 & 69.5-& 67.9 & 72.8 & 72.1 & 69.5+& 70.1\\
%70.2 & 69.1 & 68.9 & 68.3 & 74.9 & 68.4 & 69.1 & 66.6\\
%71.1 & 69.2 & 71.2 & 68.9 & 70.9 & 70.6 & 69.9 & 69.9\\
%69.4 & 69.5+& 68.5+& 70.9 & 71.6 & 68.9 & 72.0 & 70.3\\
%68.6 & 68.5-& 67.8 & 72.2 & 68.7 & 70.6 & 66.9 & 69.3\\
%71.4 & 68.7 & 74.2 & 68.8 & 71.4 & 71.8 & 67.5-& 70.4\\
%71.4 & 67.4 & 69.5-& 72.4 & 70.4 & 69.3 & 68.2 & 67.0\\
71.7 & 70.5+& 72.5-& 68.2 & 67.6 & 68.6 & 70.5-& 65.8
\vspace*{-5mm}
\end{tabular}
\end{center}
\end{table}

Mathematical formulas should be enumerated as far as they are
referenced. E.g.~Equation~(\ref{form3.1}) shows the density ...
\begin{equation}\label{form3.1}
f(x) =
\frac1{\sqrt{2\pi}\sigma}e^{-\frac{(x-\mu)^2}{2\sigma^2}}\,.
\end{equation}

Cited literature should be summarized at the end of the article in
form of a non-enumerated section. \cite{leisch02} is a reference
to an article. If the citation is in parentheses, it
should look like \citep{leisch02}.
References should be in title case.

Correspondence addresses of author(s) should be added at the end
of the manuscript.

\section{Submission and refereeing}

Consideration of the previously stated rules are of great help for
the editing process. Whenever technically possible, manuscripts
should be submitted through the editorial system at \href{www.ajs.or.at}{www.ajs.or.at}.

All contributions will be anonymously refereed. Cited literature which is
hardly available should accompany the submitted manuscript. It
should also be considered to place used and analyzed data at
disposal for the referees (if there are no legal or technical
arguments against).

Editor and referees must trust that the contribution has not been
submitted for publication at the same time at another place. It is
fair that the submitting author notifies if an earlier version has
already been submitted somewhere before.

Manuscripts stay with the publisher and referees. The refereeing
and publishing in the Austrian Journal of Statistics is free of
charge. The publisher, the Austrian Statistical Society requires a
grant of copyright from authors in order to effectively publish
and distribute this journal worldwide.

>>>>

\section{Introduction}
A brief overview of what the reader can expect of this article. What is the aim of this article and where can it be located in science. Discussion of the reproducible research movement and where the package may be beneficial. Packages like Sweave, knitr, R-markdown. How can simulation studies (the \proglang{R}-code) be published? As script? As stand-alone-package? How can ideas for specific data generation tools be shared amongst the small area community and thus be more transparent: saeSim (far-fetched but anyway).

@@

\section{Simulation studies in small area estimation}
The role of simulation studies in the field. Separation into model-based and design-based simulation studies. How does \proglang{saeSim} address these 2 perspectives? Identifying steps/phases/components of simulation studies (data generation, sampling, computation).

\section{saeSim: A simulation framework}
Introducing a simulation study as a data manipulation process. From data generation to estimating models. How can this process be mapped into the \proglang{R}-language. Present the simulation phases as flow-diagram and map the function names to the problem domain. Defining interfaces between phases: \proglang{data.frame} in, \proglang{data.frame} out. What is the difference between a design-based (fixed-population) and model-based (random-population) approach.

Why is this package helping in terms of literate programming (code is written for humans not machines) and reproducible research? Package naming conventions. Use the \proglang{\%>\%} operator to compose simulation set-ups. Reuse defined scenarios to compose new scenarios: What is a contamination scenario? A standard scenario plus contamination:

<<eval=FALSE>>=
contaminatedSetup <- standardSetup %>% sim_gen_cont()
@

\subsection{Data generation}
This is not a data generation tool. However, it supports some useful functions to add random variates to the data. Definition of outliers, see Rao (?) -- how is it addressed in the package. More than linear models, define any response with:
<<eval=FALSE>>=
setup %>% sim_resp_eq(y = g(2 * x) + e)
@

\subsection{Drawing samples}
Is present in unit-level simulation studies. Draw with simple random sampling from the whole population or within cluster/domains. Specify the sample size as integer or fraction and add weights if necessary. Wrappers around \proglang{dplyr::sample\_n} and \proglang{dplyr::sample\_frac}.

\subsection{Adding customized computations}
Adding user specified functions to the simulation process is what separates this package from a mere data generation tool. The interface is simple: Add functions with one argument which is the data at that moment and which return the modified data. I am not sure if this and the sampling section is necessary -- maybe it should just be summarized in the overview section...

\subsection{Separation of designing and running simulations}
First develop a scenario outside any looping structures, then run it R-times:

<<eval=FALSE>>=
simulationResults <- simSetup %>% sim(R = 500)
@

There is an easy back-end to run simulations in parallel (in Windows with special care):

<<eval=FALSE>>=
simulationResults <- simSetup %>% sim(R = 500, parallel = TRUE)
@

\section{Outlook}
Use this package to share and publish simulation studies alongside papers. Contribute to the package to make your ideas available. Contribute to the package and make your whole simulation study available.

### 2014-10-10 saeSim.tex
saeSim paper template

% Loading (and installing) necessary packages.
<<echo=FALSE>>=
if(!suppressPackageStartupMessages(require(saeSim))) {
  if(!require(devtools)) {
    install.packages("devtools", repos = "http://cran.at.r-project.org/", dependencies = TRUE)
    library(devtools)
  }

  install.packages(c("dplyr", "ggplot2", "MASS", "spdep"), 
                   repos = "http://cran.at.r-project.org/", dependencies = TRUE)
  install_github("wahani/saeSim")
}
@

@@

<<eval=FALSE>>=
contaminatedSetup <- standardSetup %>% sim_gen_cont()
@

\subsection{Data generation}
This is not a data generation tool. However, it supports some useful functions to add random variates to the data. Definition of outliers, see Rao (?) -- how is it addressed in the package. More than linear models, define any response with:
<<eval=FALSE>>=
setup %>% sim_resp_eq(y = g(2 * x) + e)
@

@@

<<eval=FALSE>>=
simulationResults <- simSetup %>% sim(R = 500)
@

@@

<<eval=FALSE>>=
simulationResults <- simSetup %>% sim(R = 500, parallel = TRUE)
@

>>>>

% Loading (and installing) necessary packages.

@@

\begin{Schunk}
\begin{Sinput}
> contaminatedSetup <- standardSetup %>% sim_gen_cont()
\end{Sinput}
\end{Schunk}

\subsection{Data generation}
This is not a data generation tool. However, it supports some useful functions to add random variates to the data. Definition of outliers, see Rao (?) -- how is it addressed in the package. More than linear models, define any response with:
\begin{Schunk}
\begin{Sinput}
> setup %>% sim_resp_eq(y = g(2 * x) + e)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> simulationResults <- simSetup %>% sim(R = 500)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> simulationResults <- simSetup %>% sim(R = 500, parallel = TRUE, mc.cores = 8)
\end{Sinput}
\end{Schunk}
